---
title: "Step-by-step"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step-by-step}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}

```

### First you need model(s)

Five models are provided with the MBBE package. These models were selected using [pyDarwin](https://certara.github.io/pyDarwin/html/index.html). The data source for these model is

?????

Two formatted data set is provide here (), data_seq.csv and data_sim.csv. Note that there two data sets are identical. In general, the analysis data set, for bootstrap, will not be the same as the simulation data. For example, only sparse data may be available for analysis (e.g., clinical scenarios were few samples can be collected), or a cross over study may not be feasible (e.g., long acting inflatables). Still the simulation must be done using a traditional study design, e.g., 4 period, cross over study. The typical scenario is to use a different study design that the study(ies) that generated the observed data. Had the observed data were from an adequate BE study, those data would be used to assess BE and MBBE would not be needed.

"simulation_data_path": "U:/mbbe/mbbe/inst/examples/data_sim.csv",

Several notes on the data:

1.  Data item for formulation (1,2 are reference, 3,4 are test) is include both in the analysis data set and the simulation data set.

2.  Data item for period/OCC is provided (1-4). In general, between occasion variability (e.g., by study period) should be assessed and included in the model(s).

3.  Data item for sequence (SEQ) is provided (1-2). Presumable period/OCC and sequence were tested in the model selection and not found to be supported. However, they are required for the Monte Carlo simulation and statistical testing.
